---
title: "Machine Learning Hw3"
author: "Ekta Chaudhary"
date: "10/04/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This questions will be answered using the Weekly data set, which is part of the ISLR package. This data is similar in nature to the Smarket data on the textbook (ISL, Chapter 4.6) except that it contains 1,089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010. A description of the data can be found by typing ?Weekly in the Console.

```{r}
library(ISLR)
library(MASS)
library(caret)
library(glmnet)
library(MASS)
library(e1071)
library(mlbench)
library(pROC)
library(AppliedPredictiveModeling)
```

# (a) Produce some graphical summaries of the Weekly data.

```{r}
data(Weekly)

Weekly = Weekly[,-8]

featurePlot(x = Weekly[, 1:7], 
            y = Weekly$Direction,
            scales = list(x = list(relation = "free"), 
                        y = list(relation = "free")),
            plot = "density", pch = "|", 
            auto.key = list(columns = 2))
```

# (b) Use the full data set to perform a logistic regression with Direction as the response and the five Lag variables plus Volume as predictors. Do any of the predictors appear to be statistically significant? If so, which ones?

```{r}
Weekly_dat = Weekly[,-1] #excluding year 
```

```{r}
glm.fit <- glm(Direction~., 
               data = Weekly_dat, 
               family = binomial)

summary(glm.fit)
```
Lookign at the p-values, we can say that at 5% level of significance, Lag2 is statistically significant. 